---
title: "Denver Nuggets Pregame vs. Phoenix Suns"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: scroll
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(lubridate)
library(gganimate)
library(nbastatR)
library(grid)
library(hexbin)
library(dplyr)
library(broom)
library(teamcolors)
library(knitr)
library(kableExtra)
library(plotly)
theme_set(theme_bw())
```

```{r, warning = F, message = F, include=FALSE}
denGames <- dplyr::bind_rows(
  nbastatR::teams_shots(teams = "Denver Nuggets", seasons = 2020),
  nbastatR::teams_shots(teams = "Denver Nuggets", seasons = 2021))

phxGames <- dplyr::bind_rows(
  nbastatR::teams_shots(teams = "Phoenix Suns", seasons = 2020),
  nbastatR::teams_shots(teams = "Phoenix Suns", seasons = 2021))

court <- grid::rasterGrob(jpeg::readJPEG("/Users/slaterpodgorny/Desktop/nba_court.jpeg"), 
                          width = unit(1, "npc"),
                          height = unit(1, "npc"))

games <- readRDS("/Users/slaterpodgorny/Desktop/Sports Stats/nba-games-2021-Updated1.rds")
ff <- readRDS("~/Desktop/Sports Stats/four-factors-team-end-2021.rds")
advanced <- readRDS("~/Desktop/Sports Stats/advanced-team-end-2021.rds")

nugsGames <- games %>% 
  filter(slugTeam=="DEN")
opponent <- "PHX"
prevMatchups<- nugsGames %>% 
  filter(., slugOpponent == opponent)
#select(prevMatchups, idGame)

gameIDs <- prevMatchups %>% 
  select(., idGame)

denPlayers <- nbastatR::bref_players_stats(seasons = 2021) %>% 
  filter(slugTeamBREF == "DEN")

phxPlayers <- nbastatR::bref_players_stats(seasons = 2021) %>% 
  filter(slugTeamBREF == "PHO")
```

Stat Leaders {data-orientation=columns}
=======================================================================
Stat leaders{data-width=650}
-----------------------------------------------------------------------

### Stat Leaders
Points
```{r}
#Nuggets basic Stats
denStats <- denPlayers %>% 
  mutate(ptsPerGame = ptsTotals/countGames, rbsPerGame = trbTotals/countGames, astPerGame = astTotals/countGames) %>% 
  dplyr::arrange(desc(ptsPerGame)) %>% 
  summarize(., team = slugTeamBREF, player = namePlayer, position = groupPosition, points= round(ptsPerGame,2)) #, round(rbsPerGame,2), round(astPerGame,2))
#Suns basic Stats
phxStats <- phxPlayers %>% 
  mutate(ptsPerGame = ptsTotals/countGames, rbsPerGame = trbTotals/countGames, astPerGame = astTotals/countGames) %>% 
  dplyr::arrange(desc(ptsPerGame)) %>% 
  summarize(., team = slugTeamBREF, player = namePlayer, position = groupPosition, points= round(ptsPerGame,2))

scoringleaders<- bind_rows(head(denStats, 1), head(phxStats,1))

scoringleaders%>%
  kbl() %>%
  kable_material(c("striped", "hover"))
```

Rebounds
```{r}
rbsNugs <- denPlayers %>% 
  mutate(ptsPerGame = ptsTotals/countGames, rbsPerGame = trbTotals/countGames, astPerGame = astTotals/countGames) %>% 
  dplyr::arrange(desc(ptsPerGame)) %>% 
  summarize(., team = slugTeamBREF, player = namePlayer, position = groupPosition, rebounds= round(rbsPerGame,2))

rbsSuns<- phxPlayers %>% 
  filter(namePlayer == "Deandre Ayton") %>% 
  mutate(ptsPerGame = ptsTotals/countGames, rbsPerGame = trbTotals/countGames, astPerGame = astTotals/countGames) %>% 
  dplyr::arrange(desc(ptsPerGame)) %>% 
  summarize(., team = slugTeamBREF, player = namePlayer, position = groupPosition, rebounds= round(rbsPerGame,2))  
  

rbsleaders<- bind_rows(head(rbsNugs, 1), head(rbsSuns,1))

rbsleaders%>%
  kbl() %>%
  kable_material(c("striped", "hover"))

```

Assists
```{r}
asNugs <- denPlayers %>% 
  mutate(ptsPerGame = ptsTotals/countGames, rbsPerGame = trbTotals/countGames, astPerGame = astTotals/countGames) %>% 
  dplyr::arrange(desc(astPerGame)) %>% 
  summarize(., team = slugTeamBREF, player = namePlayer, position = groupPosition, assists = round(astPerGame,2))

asSuns<- phxPlayers %>% 
  #filter(namePlayer == "Deandre Ayton") %>% 
  mutate(ptsPerGame = ptsTotals/countGames, rbsPerGame = trbTotals/countGames, astPerGame = astTotals/countGames) %>% 
  dplyr::arrange(desc(astPerGame)) %>% 
  summarize(., team = slugTeamBREF, player = namePlayer, position = groupPosition, assists=round(astPerGame,2))  
  

astleaders<- bind_rows(head(asNugs, 1), head(asSuns,1))

astleaders%>%
  kbl() %>%
  kable_material(c("striped", "hover"))
```



Column {data-width=350}
-----------------------------------------------------------------------

### Nikola Jokic Shot chart 

```{r}
## Jokic Shot Chart 
denplayer <- "Nikola Jokic"
p <- ggplot(data = filter(denGames, namePlayer == denplayer),
            aes(x = -locationX, y = locationY, col = zoneBasic))
p + annotation_custom(court, -250, 250, -50, 420) +
  geom_point(alpha = .25) +
  scale_color_brewer("Zone", palette = "Dark2") +
  coord_fixed() +
  scale_x_continuous("", breaks = NULL, limits = c(-250, 250)) +
  scale_y_continuous("", breaks = NULL, limits = c(-50, 420)) 
```

### Devin Booker Shot Chart 

```{r}
sunsplayer <- "Devin Booker"
p <- ggplot(data = filter(phxGames, namePlayer == sunsplayer),
            aes(x = -locationX, y = locationY, col = zoneBasic))
p + annotation_custom(court, -250, 250, -50, 420) +
  geom_point(alpha = .25) +
  scale_color_brewer("Zone", palette = "Dark2") +
  coord_fixed() +
  scale_x_continuous("", breaks = NULL, limits = c(-250, 250)) +
  scale_y_continuous("", breaks = NULL, limits = c(-50, 420)) 
```


Effective Field Goal {data-navmenu="Four Factors"}
======================================================================= 
### **Effective Field Goal Percentage:** The Denver Nuggets season average Effective Field Goal Percentage of 55.88 falls behind the Suns rating of 56.53.
```{r, message =F, warning=F}
df_merge <- merge(games, advanced)
df_merge<- merge(df_merge, ff) %>% 
  dplyr::mutate(., ftr = ftaTeam/fgaTeam)

currTeamsEFG <- advanced %>% 
  filter(., slugTeam %in% c("PHX", "DEN")) %>% 
  group_by(slugTeam) %>% 
  summarize(., pctEFG = mean(pctEFG))


p <- ggplot(data = advanced,
            aes(x = pctEFG))
p + geom_density() +
  geom_vline(data = currTeamsEFG,
             aes(xintercept = pctEFG, col = slugTeam)) +
  scale_color_brewer("team", palette = "Set1") +
  labs(x = "Efective Field Goal Percentage", 
       y = "density") +
  theme_bw()
```
 

Offensive Rebounds {data-navmenu="Four Factors"}
======================================================================= 
### **Offensive Rebounds:** The Denver Nuggets season average Offensive Rebound Percentage of 28.86 exceeds the Suns rating of 24.57.
```{r}
currTeamsOREB <- advanced %>% 
  filter(., slugTeam %in% c("PHX", "DEN")) %>% 
  group_by(slugTeam) %>% 
  summarize(., pctOREB = mean(pctOREB))


p <- ggplot(data = advanced,
            aes(x = pctOREB))
p + geom_density() +
  geom_vline(data = currTeamsOREB,
             aes(xintercept = pctOREB, col = slugTeam)) +
  scale_color_brewer("team", palette = "Set1") +
    labs(x = "Offensive Rebound Percentage", 
       y = "density") +
  theme_bw()
```


Turnovers {data-navmenu="Four Factors"}
======================================================================= 
### **Turnover Percentage:** The Denver Nuggets season average Turnover Percentage of 13.58 exceeds the Suns rating of 12.57.
```{r}
currTeamsTOV<- advanced %>% 
  filter(., slugTeam %in% c("PHX", "DEN")) %>% 
  group_by(slugTeam) %>% 
  summarize(., pctTOV=mean(pctTOVTeam))


p <- ggplot(data = advanced,
            aes(x = pctTOVTeam))
p + geom_density() +
  geom_vline(data = currTeamsTOV,
             aes(xintercept = pctTOV, col = slugTeam)) +
  scale_color_brewer("team", palette = "Set1") +
    labs(x = "Turnover Percentage", 
       y = "density") +
  theme_bw()
```

Free Throw Attempts{data-navmenu="Four Factors"}
======================================================================= 
### **Free Throw Rate:** The Denver Nuggets season average Free Throw Rate of 22.1 exceeds the Suns rating of 21.67.
```{r}
currTeamsFTR<- df_merge %>% 
  filter(., slugTeam %in% c("PHX", "DEN")) %>% 
  group_by(slugTeam) %>% 
  summarize(., FTR = mean(ftr))

p <- ggplot(data = df_merge,
            aes(x = ftr))
p + geom_density() +
  geom_vline(data = currTeamsFTR,
             aes(xintercept = FTR , col = slugTeam)) +
  scale_color_brewer("team", palette = "Set1") +
    labs(x = "Free Throw Rate", 
       y = "density") +
  theme_bw()
```

Four Factors {data-navmenu="Previous Matchups"}
=======================================================================

Column {data-width=650}
-----------------------------------------------------------------------

#### Effective Field Goal Percentage
```{r}
#Wins 
prevDataFF <- merge(prevMatchups,df_merge) %>% 
  dplyr::select(., pctEFG, pctOREB, pctTOVTeam, rateFTA, outcomeGame) #%>% 
  #dplyr::mutate(.,pctTOVTeam = pctTOVTeam/100)

p <- ggplot(data = advanced,
            aes(x = pctEFG))
p + geom_density() +
  geom_vline(data = prevDataFF,
             aes(xintercept = pctEFG, col = outcomeGame)) +
  scale_color_brewer("team", palette = "Set1")  +
    labs(x = "Effective Field Goal Percentage", 
       y = "density") +
  theme_bw()


```


#### Turnover Percentage
```{r}
p <- ggplot(data = advanced,
            aes(x = pctTOVTeam))
p + geom_density() +
  geom_vline(data = prevDataFF,
             aes(xintercept = pctTOVTeam, col = outcomeGame)) +
  scale_color_brewer("team", palette = "Set1")  +
    labs(x = "Turnover Percentage", 
       y = "density") +
  theme_bw()
```

Column {data-width=650}
-----------------------------------------------------------------------

#### Offensive Rebounding Percentage
```{r}
p <- ggplot(data = advanced,
            aes(x = pctOREB))
p + geom_density() +
  geom_vline(data = prevDataFF,
             aes(xintercept = pctOREB, col = outcomeGame)) +
  scale_color_brewer("team", palette = "Set1") +
    labs(x = "Offensive Rebounding Percentage", 
       y = "density") +
  theme_bw()
```


#### Freethrow Rate
```{r}
p <- ggplot(data = df_merge,
            aes(x = ftr))
p + geom_density() +
  geom_vline(data = prevDataFF,
             aes(xintercept = rateFTA, col = outcomeGame)) +
  scale_color_brewer("team", palette = "Set1") +
    labs(x = "Free Throw Rate", 
       y = "density") +
  theme_bw()
```

Box Scores{data-navmenu="Previous Matchups"}
=======================================================================

Column {data-width=250}
-----------------------------------------------------------------------

```{r, message =F, warning=F, include =F }
game1<-box_scores(game_ids = 	22000076, box_score_types = "Traditional", result_types="player")
denboxgame1<-(game1$dataBoxScore[[1]]) %>% 
  filter(., slugTeam == "DEN") %>% 
  summarize(., player = namePlayer, points= pts, rebounds = treb, assists = ast)

```
#### Game 1: Nuggets
```{r}
denboxgame1%>%
  kbl() %>%
  kable_material(c("striped", "hover"))
```

Column {data-width=250}
-----------------------------------------------------------------------

```{r, message =F, warning=F, include = F}
game1<-box_scores(game_ids = 	22000076, box_score_types = "Traditional", result_types="player")
phxboxgame1<-(game1$dataBoxScore[[1]]) %>% 
  filter(., slugTeam == "PHX") %>% 
  summarize(., player = namePlayer, points= pts, rebounds = treb, assists = ast)


```

#### Game 1: Suns
```{r}
phxboxgame1%>%
  kbl() %>%
  kable_material(c("striped", "hover"))
```

Shot Charts{data-navmenu="Previous Matchups"}
=======================================================================


Column {data-width=300}
-----------------------------------------------------------------------

Game 1 - Loss
```{r}
game1Id<-22000076
p <- ggplot(data = filter(denGames, idGame == game1Id),
            aes(x = -locationX, y = locationY,
                col = typeEvent))
p + annotation_custom(court, -250, 250, -50, 420) +
  geom_point(alpha = .5) +
  scale_color_brewer("Zone", labels = c("Made", "Miss"), palette = "Set1") +
  coord_fixed() +
  scale_x_continuous("", breaks = NULL, limits = c(-250, 250)) +
  scale_y_continuous("", breaks = NULL, limits = c(-50, 420))

fieldgoalsg1<- nugsGames %>% 
  filter(idGame ==game1Id)
  




```

#### Game 1

```{r}
select(fieldgoalsg1, fgmTeam, fgaTeam, pctFGTeam) %>% 
  kbl() %>%
  kable_material(c("striped", "hover"))
select(fieldgoalsg1, fg3mTeam, fg3aTeam, pctFG3Team) %>% 
  kbl() %>%
  kable_material(c("striped", "hover"))
```

Column {data-width=300}
-----------------------------------------------------------------------

Game 2 - Win
```{r}
game2Id<-22000238
p <- ggplot(data = filter(denGames, idGame == game2Id),
            aes(x = -locationX, y = locationY,
                col = typeEvent))
p + annotation_custom(court, -250, 250, -50, 420) +
  geom_point(alpha = .5) +
  scale_color_brewer("Zone", labels = c("Made", "Miss"), palette = "Set1") +
  coord_fixed() +
  scale_x_continuous("", breaks = NULL, limits = c(-250, 250)) +
  scale_y_continuous("", breaks = NULL, limits = c(-50, 420))

fieldgoalsg2<- nugsGames %>% 
  filter(idGame ==game2Id)


```

#### Game 2
```{r}
select(fieldgoalsg2, fgmTeam, fgaTeam, pctFGTeam) %>% 
  kbl() %>%
  kable_material(c("striped", "hover"))
select(fieldgoalsg2, fg3mTeam, fg3aTeam, pctFG3Team) %>% 
  kbl() %>%
  kable_material(c("striped", "hover"))
```

Column {data-width=300}
-----------------------------------------------------------------------

Game 3 - Win 
```{r}
game3Id<-22000248
p <- ggplot(data = filter(denGames, idGame == game3Id),
            aes(x = -locationX, y = locationY,
                col = typeEvent))
p + annotation_custom(court, -250, 250, -50, 420) +
  geom_point(alpha = .5) +
  scale_color_brewer("Zone", labels = c("Made", "Miss"), palette = "Set1") +
  coord_fixed() +
  scale_x_continuous("", breaks = NULL, limits = c(-250, 250)) +
  scale_y_continuous("", breaks = NULL, limits = c(-50, 420))

fieldgoalsg3<- nugsGames %>% 
  filter(idGame ==game3Id)


```

#### Game 3
```{r}
select(fieldgoalsg3, fgmTeam, fgaTeam, pctFGTeam) %>% 
  kbl() %>%
  kable_material(c("striped", "hover"))
select(fieldgoalsg3, fg3mTeam, fg3aTeam, pctFG3Team) %>% 
  kbl() %>%
  kable_material(c("striped", "hover"))
```

Offensive Keys{data-navmenu="Keys"}
=======================================================================

Column {data-width=650}
-----------------------------------------------------------------------
### *Keys to the Game:*
#### Offensive Keys 
- Share the ball
- Hit the glass
- Get Porter going
- Balance the attack 

```{r, warning= F, message = F, include = F}
porterWins <- game_logs(seasons = 2021, results_types = "player") %>% 
  filter(slugTeam == "DEN",namePlayer =="Michael Porter Jr.", outcomeGame == "W") %>% 
  summarize(., points= round(mean(pts),2), assists = round(mean(ast),2), rebounds = round(mean(treb),2))

porterLs <- game_logs(seasons = 2021, results_types = "player") %>% 
  filter(namePlayer =="Michael Porter Jr.", outcomeGame == "L") %>% 
  summarize(., points= round(mean(pts),2), assists = round(mean(ast),2), rebounds = round(mean(treb),2))
```

#### Assists vs Point scored

```{r}

p <- ggplot(data = nugsGames,
            aes(x = ptsTeam, y = astTeam,
                col = outcomeGame, text = paste('Result: ', outcomeGame, '\n',
                             'Points: ', ptsTeam, '\n')))+
  geom_point(alpha = .5)  +
    labs(x = "Points", 
       y = "Assists") +
  theme_bw()

ggplotly(p, tooltip ="Text")

```

Column {data-width=650}
-----------------------------------------------------------------------

Denver Averages vs Denver Wins 
```{r}
denAverages <- df_merge %>% 
  filter(slugTeam=="DEN") %>% 
  summarize(., points= round(mean(ptsTeam),2), assists = round(mean(astTeam),2), rebounds = round(mean(trebTeam),2),pctEFG = round(mean(pctEFG),2),pctOREB = round(mean(pctOREB),2), pctTOV=round(mean(pctTOVTeam),2), FTR = round(mean(ftr),2))

denWins <- df_merge %>% 
  filter(slugTeam=="DEN", outcomeGame =="W") %>% 
  summarize(., points= round(mean(ptsTeam),2), assists = round(mean(astTeam),2), rebounds = round(mean(trebTeam),2),pctEFG = round(mean(pctEFG),2),pctOREB = round(mean(pctOREB),2), pctTOV=round(mean(pctTOVTeam),2), FTR = round(mean(ftr),2))


denAverages%>%
  kbl() %>%
  kable_material(c("striped", "hover"))

denWins%>%
  kbl() %>%
  kable_material(c("striped", "hover"))
```

#### Offensive Rating and Offensive Rebound Percentage

```{r}
p <- ggplot(data = (df_merge %>% 
                      filter(slugTeam =="DEN")),
            aes(x = pctOREB, y = ortg,
                col = outcomeGame, text = paste('Result: ', outcomeGame, '\n',
                             'Points: ', ptsTeam, '\n')))+
  geom_point(alpha = .5)  +
    labs(x = "Offensive Rebound Percentage", 
       y = "Offensive Rating") +
  theme_bw()

ggplotly(p, tooltip ="Text")
```

#### Michael Porter Jr when the Nuggets win vs when they lose
```{r}
porterWins %>% 
  kbl() %>%
  kable_material(c("striped", "hover"))
porterLs %>% 
  kbl() %>%
  kable_material(c("striped", "hover"))
```

Defensive Keys{data-navmenu="Keys"}
=======================================================================

Column {data-width=650}
-----------------------------------------------------------------------

#### Defensive Keys 
- Contain Booker: Scored 30 or more points in 22 games, the Suns won 15 of those games
- Shut down Bridges: most change in scoring between wins and losses among top scorers
- Defend the perimeter: Suns shoot about 7% worse from 3 when they lose

```{r, warning = F, message = F, include = F}
bridgesWins <- game_logs(seasons = 2021, results_types = "player") %>% 
  filter(slugTeam == "PHX", namePlayer =="Mikal Bridges", outcomeGame == "W") %>% 
  summarize(., points= round(mean(pts),2), assists = round(mean(ast),2), rebounds = round(mean(treb),2))

bridgesLs <- game_logs(seasons = 2021, results_types = "player") %>% 
  filter(slugTeam == "PHX", namePlayer =="Mikal Bridges", outcomeGame == "L") %>% 
  summarize(., points= round(mean(pts),2), assists = round(mean(ast),2), rebounds = round(mean(treb),2))

booker <- game_logs(seasons = 2021, results_types = "player") %>% 
  filter(slugTeam == "PHX", namePlayer =="Devin Booker", pts >= 30) 

phxWins <- df_merge %>% 
  filter(slugTeam == "PHX", outcomeGame == "W") %>% 
  summarize(., fg3PCT = mean(pctFG3Team))

phxLs <- df_merge %>% 
  filter(slugTeam == "PHX", outcomeGame == "L") %>% 
  summarize(., fg3PCT = mean(pctFG3Team))

phxStatsplot <-df_merge %>% 
  filter(slugTeam == "PHX")
 
```

```{r}
p <- ggplot(data = phxStatsplot,
            aes(x = ptsTeam, y = pctFG3Team,
                col = outcomeGame, text = paste('Result: ', outcomeGame, '\n',
                             'Points: ', ptsTeam, '\n')))+
  geom_point(alpha = .5)  +
    labs(x = "Points", 
       y = "3 point percentage") +
  theme_bw()

ggplotly(p, tooltip ="Text")

```

Mikal Bridges when the Suns win vs when they lose

```{r}

bridgesWins%>%
  kbl() %>%
  kable_material(c("striped", "hover"))

bridgesLs%>%
  kbl() %>%
  kable_material(c("striped", "hover"))
```

Column {data-width=650}
-----------------------------------------------------------------------

### Suns when Booker scores 30 or more

```{r}
p<-ggplot(data = booker,
            aes(x = outcomeGame, fill = outcomeGame))
p+geom_bar()+
  scale_fill_manual(values=c("#636363", "#e6550d"))

```






